---
title: "week8-section"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(sf)
library(terra)
library(dismo)
library(tmap)
```

```{r}
magpie <- read_csv(here::here("data", "data", "magpie_obvs.csv"))
tule_elk <- read_csv(here::here("data", "data", "tule_elk_obvs.csv"))

bioclim_dir <- here::here("data", "data", "climate", "wc2.1_2.5m")
bioclim <- list.files(bioclim_dir, pattern = glob2rx("*.tif$"), full.names = TRUE)
bioclim_rast <- rast(bioclim)
```

Update the bioclim_rast layer names to match the Bioclimatic variables using names()

```{r}
names(bioclim_rast) <- c("annualMeanTemp", "meanDiurnalRange", "isothermality", "tempSeasonality", "maxTempWarmMonth", "maxTempColdMonth", "tempAnnualRange", "meanTempWetQ", "meanTempDryQ", "meanTempWarmQ", "meanTempColdQ", "annualPrecip", "precipWetMonth", "precipDryMonth", "precipSeasonality", "precipWetQ", "precipDryQ", "precipWarmQ", "precipColdQ")

```

Use st_as_sf to convert and find the geographic extent of species occurrence data magpie using st_bbox()

```{r}
magpie_sf <- magpie %>%
  rename(long = longitude,
         lat = latitude) %>%
  drop_na(long, lat) %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326)


```

Crop bioclim_rast to match the extent of species occurrence data magpie

```{r}
bioclim_crop <- crop(bioclim_rast, magpie_sf)

```

Extract points from the cropped bioclim_rast for each occurrence in magpie

```{r}
bioclim_mask <- mask(bioclim_crop, magpie_sf) 
```

```{r}
plot(bioclim_crop$annualMeanTemp)
plot(bioclim_mask$annualMeanTemp)
```

```{r}
# Extract points from raster for all species occurrences
bioClim_pts <- as_tibble(extract(bioclim_crop, magpie_sf))
```

Create “background values” by using dismo::randomPoints() to generate random sample points from the cropped bioclim_rast

```{r}
set.seed(42) # for reproducibility
# Generate random sample points from raster
random_pts <- dismo::randomPoints(mask = raster(bioclim_crop[["annualMeanTemp"]]),
                                  n = nrow(magpie) * 2,
                                  ext = magpie_sf)

```

Extract points from bioclim_rast for each random sample point generated in Step #5

```{r}
# Extract points from raster for random sample points
bioClim_random_pts <- as_tibble(extract(bioclim_crop, random_pts))
```

Plot species climate niche and background climate (temperature vs. precipitation)

```{r}
map_1 <- tm_shape(raster(bioclim_crop[["annualPrecip"]])) +
  tm_raster(palette = "Blues", title = "Annual Precipitation") +
  tm_shape(magpie_sf) +
  tm_dots(col = "#3a5a40", size = 0.15) +
  tm_layout(legend.position = c("left", "bottom"),
            legend.bg.color = "white")

map_2 <- tm_shape(raster(bioclim_crop[["annualMeanTemp"]])) +
  tm_raster(palette = "-RdYlBu", title = "Annual Mean Temp") +
  tm_shape(magpie_sf) +
  tm_dots(col = "#3a5a40", size = 0.15) +
  tm_layout(legend.position = c("left", "bottom"),
            legend.bg.color = "white")

tmap_arrange(map_1, map_2)
```


```{r}

```

